Recruits, abundance, and yield time series and contours

### 2023-11-07 ms_figures_supp_figs.qmd
refactor and revise plot functions for main figures, data prep steps interim commit

waa_g <- sim_species_parms[[nice_species]][["biom_const"]] * sim_species_derived_vars[[nice_species]][["length_at_age"]] ^
                    sim_species_parms[[nice_species]][["biom_exp"]]

# use functions to create useful data structures
# FLEP to F reference tables by species
spec_fleps_2_fs <-
  get_fs_from_fleps(species = nice_species,
                    flep_inds = species_flep_fished_ind,
                    f_vals =  fishing_mortality_values)

@@

set.seed(11)
num_sims_2_avg <- 5
sims_2_use <- sort(sample(1:sim_nums, size = num_sims_2_avg, replace = FALSE))

table_name <- "no_dispersal_blue_rockfish_mei_135_75"

@@

out_2 <- out %>%
  group_by(sim_num, reserve_frac, flep_ratio) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm=TRUE)) %>%
  ungroup()

ggplot(out_2, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(flep_ratio),
                z = abundance)) +
  facet_grid(sim_num ~ .) +
  geom_contour_filled()

out_3 <- out_2 %>%
  group_by(reserve_frac, flep_ratio) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm = TRUE)) %>%
  ungroup()

ggplot(out_3, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(flep_ratio),
                z = abundance)) +
  geom_contour_filled()

>>>>

# waa_g <- sim_species_parms[[nice_species]][["biom_const"]] * sim_species_derived_vars[[nice_species]][["length_at_age"]] ^
#                     sim_species_parms[[nice_species]][["biom_exp"]]
# 
# 
# # use functions to create useful data structures
# # FLEP to F reference tables by species
# spec_fleps_2_fs <-
#   get_fs_from_fleps(species = nice_species,
#                     flep_inds = species_flep_fished_ind,
#                     f_vals =  fishing_mortality_values) 
#

@@

## Species parms

brf_waa_g <- sim_species_parms[["Blue rockfish"]][["biom_const"]] * sim_species_derived_vars[["Blue rockfish"]][["length_at_age"]] ^
                    sim_species_parms[["Blue rockfish"]][["biom_exp"]]

brf_waa_g_df <- data.frame(age = 1: sim_species_parms[["Blue rockfish"]]$A_max,
                           waa_g = brf_waa_g)

# use functions to create useful data structures
# FLEP to F reference tables by species
spec_fleps_2_fs <-
  get_fs_from_fleps(species = "Blue rockfish",
                    flep_inds = species_flep_fished_ind,
                    f_vals =  fishing_mortality_values)

## Qry 
set.seed(11)
num_sims_2_avg <- 5
sims_2_use <- sort(sample(1:sim_nums, size = num_sims_2_avg, replace = FALSE))

table_name <- "no_dispersal_blue_rockfish_white_135_75"

@@

get_f_maxes_yield <- function(out_df,
                              fleps_2_fs) {
  # Find the fishing mortality level that maximizes yield
  # with no reserves
  F_maxes_yield <- out %>%
    dplyr::filter(reserve_frac == "0") %>%
    dplyr::left_join(., y = fleps_2_fs, by = join_by(flep_ratio == flep_vals_ch)) %>%
    dplyr::group_by(f_vals, flep_ratio, year) %>%
    dplyr::summarise(yield = sum(yield),
                     abundance = sum(abundance)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(f_vals, flep_ratio) %>%
    dplyr::summarize(yield = mean(yield, na.rm = TRUE),
                     abundance = mean(abundance, na.rm = TRUE)) %>%
    ungroup() %>%
    dplyr::filter(yield == max(yield)) %>%
    .$f_vals

  return(F_maxes_yield)
}

df_add_fs_out <- function(out,
                          spec_fleps_2_fs,
                          F_maxes_yield) {
out_fs <- out %>%
  mutate(flep_ratio_ch = as.character(flep_ratio)) %>%
  dplyr::left_join(., y = spec_fleps_2_fs, by = join_by(flep_ratio == flep_vals_ch)) %>%
  dplyr::mutate(f_fmsy = round(f_vals / F_maxes_yield, 2)) %>%
  dplyr::left_join(., y = brf_waa_g_df, join_by(age == age)) %>%
  dplyr::mutate(biomass_at_age = (abundance * waa_g) / 1E6)

}

get_median_yield <- function(df_out_fs) {
  # returns the mean of median yield at at F = F_msy 
  # for a coastline with no reserves across sims

  stopifnot(colnames(df_out_fs) == c("sim_num", "reserve_frac", "flep_ratio",
                                     "year", "age", "abundance", "yield", 
                                     "flep_ratio_ch", "flep_vals", "f_vals", 
                                     "f_fmsy", "waa_g", "biomass_at_age"))

  sim_num_2_use <- min(df_out_fs$sim_num)
  med_yield <- df_out_fs %>%
    dplyr::filter(near(f_fmsy, 1), near(as.numeric(reserve_frac), 0))  %>%
    dplyr::group_by(sim_num,
                    reserve_frac,
                    f_fmsy,
                    year) %>%
    dplyr::summarise(ann_yield = sum(yield)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(sim_num) %>%
    dplyr::summarise(med_ann_yield = median(ann_yield)) %>%
    dplyr::ungroup() %>%
    dplyr::summarise(avg_med_yield_f_fmsy = mean(med_ann_yield)) %>%
    .$avg_med_yield_f_fmsy
  return(med_yield)
}

get_median_biomass <- function(df_out_fs) {
  # returns the mean of median abundance at at F = F_msy 
  # for a coastline with no reserves across sims
  stopifnot(colnames(df_out_fs) == c("sim_num", "reserve_frac", "flep_ratio",
                                     "year", "age", "abundance", "yield", 
                                     "flep_ratio_ch", "flep_vals", "f_vals", 
                                     "f_fmsy", "waa_g", "biomass_at_age"))
  med_biomass <- df_out_fs %>%
    dplyr::filter(near(f_fmsy, 1), near(as.numeric(reserve_frac), 0))  %>%
    dplyr::group_by(sim_num,
                    reserve_frac,
                    flep_ratio,
                    year) %>%
    dplyr::summarise(ann_biomass = sum(biomass_at_age)) %>%
    dplyr::ungroup() %>%
     dplyr::group_by(sim_num) %>%
    dplyr::summarise(med_ann_biomass = median(ann_biomass)) %>%
    dplyr::summarise(avg_med_biomass_f_fmsy = mean(med_ann_biomass)) %>%
    .$avg_med_biomass_f_fmsy
  return(med_biomass)
}

df_yld_bm_stats_by_sim_f_frac_year <- function(df_out_fs,
                                      sims_num_2_use) {
  stopifnot(colnames(df_out_fs) == c("sim_num", "reserve_frac", "flep_ratio",
                                     "year", "age", "abundance", "yield", 
                                     "flep_ratio_ch", "flep_vals", "f_vals", 
                                     "f_fmsy", "waa_g", "biomass_at_age"))
  # make a sub function
  out_df <- df_out_fs %>%
    dplyr::filter(near(sim_num, sim_num_2_use)) %>%
    dplyr::filter(reduce(map(
      frac_reserves,
      near,
      x = as.numeric(reserve_frac),
      tol = 1e-4
    ), `|`)) %>%
    dplyr::group_by(sim_num, reserve_frac, f_fmsy, year) %>%
    dplyr::summarize(annual_yield = sum(yield),
                     annual_biomass = sum(biomass_at_age)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(reduce(map(
      f_fmsys, near, x = f_fmsy, tol = 1e-4
    ), `|`))
  return(out_df)
}

# df_exp_spec_noise_rec_fs
prep_ann_yield_abund_for_plot <- function(df_out_fs,
                                          frac_reserves = c(0, 0.25, 0.50),
                                          f_fmsys = c(0, 0.52, 1.00, 2.09)) {

  sim_num_2_use <- min(df_out_fs$sim_num)
  if (length(f_fmsys) < 1 | is.null(f_fmsys)) {
    stop("Please specify f_fmsy values to use")
  }

  if (length(frac_reserves) < 1 | is.null(frac_reserves)) {
    stop("Please specify frac_reserves values to use")
  }

    # make a sub function
    supp_plt_df <- main_plt_df %>%
      dplyr::mutate(below_thresh_yld = if_else(annual_yield <= med_yield, 1, 0),
                    below_thresh_bm = if_else(annual_biomass <= med_biomass, 1, 0)) %>%
      dplyr::group_by(sim_num, reserve_frac, f_fmsy) %>%
      dplyr::summarise(
        mean_yield = round(mean(annual_yield, na.rm = TRUE), 2),
        mean_biomass = round(mean(annual_biomass, na.rm = TRUE), 2),
        sd_yield = round(sd(annual_yield, na.rm = TRUE), 2),
        sd_biomass = round(sd(annual_biomass, na.rm = TRUE), 2),
        cv_yield = round(sd(annual_yield, na.rm = TRUE)/mean(annual_yield, na.rm = TRUE), 2),
        cv_biomass = round(sd(annual_biomass, na.rm = TRUE)/mean(annual_biomass, na.rm = TRUE), 2),
        yrs_below_yield_thresh = round(sum(below_thresh_yld)/135, 2),
        yrs_below_biomass_thresh = round(sum(below_thresh_bm)/135, 2)
      ) %>%
      dplyr::ungroup() %>%
      dplyr::group_by(reserve_frac, f_fmsy) %>%
      dplyr::summarize(
        mean_yield = round(mean(mean_yield, na.rm = TRUE), 2),
        mean_yield_lab = paste0("Mean yield: ", mean_yield),
        mean_biomass = round(mean(mean_biomass, na.rm = TRUE), 2),
        mean_biomass_lab = paste0("Mean biomass: ", mean_yield),
        sd_yield = round(sd(annual_yield, na.rm = TRUE), 2),
        sd_yield_lab = paste0("SD yield: ", sd_yield),
        sd_abundance = round(sd(annual_abundance, na.rm = TRUE), 2),
        sd_abundance_lab = paste0("SD biomass: ", sd_yield),
        cv_yield = round(sd(annual_yield, na.rm = TRUE)/mean(annual_yield, na.rm = TRUE), 2),
        cv_yield_lab = paste0("CV yield: ", cv_yield),
        cv_biomass = round(sd(mean_abundance, na.rm = TRUE)/mean(mean_abundance, na.rm = TRUE), 2),
        cv_biomass_lab = paste0("CV biomass: ", cv_biomass),
        yrs_below_yield_thresh = paste("% YR below yield thresh: ", round(sum(below_thresh_yld)/135, 2)),
        yrs_below_biomass_thresh = paste("% YR below biomass thresh: ", round(sum(below_thresh_bm)/135, 2))
      ) %>%
      ungroup()
    return(list(main_plt_df, supp_plt_df, med_yield))
  }

F_brf_maxes_yield <- get_f_maxes_yield(out_df = out,
                              fleps_2_fs = spec_fleps_2_fs)
F_brf_maxes_yield
out_fs

med_yield <- get_median_yield(df_out_fs = df_out_fs) 
med_yield
med_biomass <- get_median_biomass(df_out_fs = df_out_fs)
med_biomass
out <- df_yld_bm_stats_by_sim_f_frac_year(df_out_fs = df_out_fs,
                                      sims_num_2_use = sims_2_use)
out
test_plot_df <- prep_ann_yield_abund_for_plot(
  df_out_fs = out_fs,
  frac_reserves = c(0, 0.25, 0.50),
  f_fmsys = c(0, 0.52, 1.00, 2.09)
)

out_2 <- out_fs %>%
  group_by(sim_num, reserve_frac, f_fmsy) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm=TRUE)) %>%
  ungroup()

out_3 <- out_2 %>%
  group_by(reserve_frac, f_fmsy) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm = TRUE)) %>%
  ungroup()

ggplot(out_2, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(f_fmsy),
                z = abundance)) +
  facet_grid(sim_num ~ .) +
  geom_contour_filled()

ggplot(out_3, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(f_fmsy),
                z = abundance)) +
  geom_contour_filled()
